---
title: "DS 201 Homework 2"
format: 
  html:
    theme: pulse  
---

This is Homework Assignment 2 for DS 201. You can view the course code for this assignment on GitHub: [view the source code](https://github.com/jmgraham30/ds_201_assignments/blob/main/homework_02/index.qmd).

For your amusement: What's orange and sounds like a parrot? A carrot!

## Problems

### Problem 1

a. Create a RStudio project on your computer and name it `homework_01_lastname` where you should replace "lastname" with your last name. For example, I would create a project named `homework_01_graham`. 

b. Add a subfolder to your project title `data`. 

c. While in your project, create a Quarto document with title "Homework 1" and add your name as the author. Save this file as `lastname_homework_01` replacing the "lastname" part with your last name.

d. Add the following R code chunk to your Quarto document:

```{r}
#| message: false
#| warning: false
#| code-fold: false
#| echo: true

# load useful libraries
library(tidyverse)
library(tidytuesdayR)
```

### Problem 2

Skim through the information in the following repository:

- Childcare Costs [view the repository](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-05-09/readme.md)

Write a paragraph in your Quarto document describing the data that is contained in this repository. Make sure that your paragraph addresses the following points:

a. The source of the data. 

b. The format the data is saved as. That is, is it saved as a csv file, a dat file, etc.?

c. A general description of what information is recorded in the data.

### Problem 3

Explore the website associated with the data source, [view the website](https://www.dol.gov/agencies/wb/topics/featured-childcare).

a. What type of research or questions are or could be addressed with the data from Problem 2? Write a paragraph answering this question in your Quarto document. 

### Problem 4

Run the following R code in a code chunk in your Quarto document:

```{r}
#| message: false
#| warning: false
#| code-fold: false
#| echo: true

# load the data
childcare_costs <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-09/childcare_costs.csv')
counties <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-05-09/counties.csv')
```

a. What do these commands do? 

b. Add an R code chunk in your Quarto document and use the `glimpse` command to examine the data you have just loaded. 

c. Use the out put from the `glimpse` command to determine the number of rows and columns that are contained in the data frames `childcare_costs` and `counties`.